---
title: Henard
author: Brett Davidheiser
date: 05/07/2019
output: html_document
version: EA code version 3, C correction only without a blank correction, Katie Snell
---
## NOTES about data reduction and run:
Fill in with notes about the run: e.g. settings for culling, and extra standards culled, which column of corrected data is used in the discrimination correction.
integrate

Some miss nameing in the run list things were offset by one.

16200__empty_0.dxf Had a large sample in it. What happend? Left from last run?

samples culled or too small for linearity range: 


Standards culled and why:
The collagen mon standard #16228 was just way off in d13C. It looked normal in all other ways.

Notes for the lab user
look for memory efect from tracer samples.


```{r}
#devtools::install_github("KopfLab/isoreader")
#devtools::install_github("cubessil/isoprocessCUBES")
```

## Load libraries, data, and define standards
load necessary R libraries and data, define accepted standard values for correcting standards
```{r load-stuff, message=FALSE,warning=FALSE}
library(grid)
library(tidyverse)
library(isoprocessCUBES) 
library(knitr)
library(stringr)
library(openxlsx)
library(plotly)
library(isoreader)
library(isoprocessor)
```

load data and add inverse area columns
```{r, message=FALSE,warning=FALSE}

rawfiles.directory<- file.path(".")
dxf_files<-iso_read_continuous_flow(rawfiles.directory,read_vendor_data_table = TRUE)

session<-"Henard run 5"
dxf_files <- iso_omit_files_with_problems(dxf_files)
rawinfo <- dxf_files %>% 
  iso_get_file_info() %>% 
  # pick favorite file info fields
  select(file_id, `Analysis`,`Row`, `Identifier 1`, `Identifier 2`,`Comment`, `Preparation`, `file_datetime`,`Method`) %>% 
  data.frame()
colnames(rawinfo)[colnames(rawinfo)=="Row"] <- "row"
colnames(rawinfo)[colnames(rawinfo)=="Identifier.1"] <- "Identifier1"
colnames(rawinfo)[colnames(rawinfo)=="Identifier.2"] <- "mass"
colnames(rawinfo)[colnames(rawinfo)=="Comment"] <- "carousel"
colnames(rawinfo)[colnames(rawinfo)=="Preparation"] <- "type"



#saveRDS(dxf_files, file = "help.rds")
getratio <- function(obs){
  ratiodata <- obs %>%
  iso_get_vendor_data_table(with_units = FALSE)%>%
  select(`Nr.`, `Start`, `End`, `Intensity 28`, `Intensity 44`, `Intensity 45`,`Ampl 45`,`d 29N2/28N2`,`d 13C/12C`)
  fileid = obs$file_info$file_id
  ratiodata %>% mutate(file_id = fileid)
}

rawratio <- bind_rows(lapply(dxf_files, getratio))
rawratio <- subset(rawratio, Start >115)
rawratio <- subset(rawratio, Start <350)
colnames(rawratio)[colnames(rawratio)=="Intensity 28"] <- "area28"
colnames(rawratio)[colnames(rawratio)=="Intensity 44"] <- "area44"
colnames(rawratio)[colnames(rawratio)=="Intensity 45"] <- "area45"
colnames(rawratio)[colnames(rawratio)=="d 13C/12C"] <- "d13C"
colnames(rawratio)[colnames(rawratio)=="d 29N2/28N2"] <- "d29N"


raw.data<-full_join(rawinfo, rawratio, by = "file_id" )
raw.data<- transform(raw.data, "row" = as.numeric(`row`))
raw.data<- transform(raw.data, "mass" = as.numeric(`mass`))
raw.data<- transform(raw.data, "Analysis" = as.numeric(`Analysis`))
raw.dataCarbon <- raw.data[!(is.na(raw.data$`d13C`) | raw.data$`d13C`==""), ]
data<-raw.dataCarbon
data$inv.area44<-1/data$area44


```
```{r samples off by one}
  ID = 18326
  data[ data$Analysis == ID , "Identifier1"] <- "30h OD 0.51"
  data[ data$Analysis == ID , "mass"] <- "16"
  data[ data$Analysis == ID , "carousel"] <- "5"
  data[ data$Analysis == ID , "type"] <- "sample"
  
  ID = 18327
  data[ data$Analysis == ID , "Identifier1"] <- "1.1"
  data[ data$Analysis == ID , "mass"] <- "25"
  data[ data$Analysis == ID , "carousel"] <- "5"
  data[ data$Analysis == ID , "type"] <- "sample"
  
  ID = 18328
  data[ data$Analysis == ID , "Identifier1"] <- "1.2"
  data[ data$Analysis == ID , "mass"] <- "18"
  data[ data$Analysis == ID , "carousel"] <- "5"
  data[ data$Analysis == ID , "type"] <- "sample"
  
  ID = 18329
  data[ data$Analysis == ID , "Identifier1"] <- "1.3"
  data[ data$Analysis == ID , "mass"] <- "35"
  data[ data$Analysis == ID , "carousel"] <- "5"
  data[ data$Analysis == ID , "type"] <- "sample"
  
  ID = 18330
  data[ data$Analysis == ID , "Identifier1"] <- "2.1"
  data[ data$Analysis == ID , "mass"] <- "20"
  data[ data$Analysis == ID , "carousel"] <- "5"
  data[ data$Analysis == ID , "type"] <- "sample"
  
  ID = 18331
  data[ data$Analysis == ID , "Identifier1"] <- "2.2"
  data[ data$Analysis == ID , "mass"] <- "29"
  data[ data$Analysis == ID , "carousel"] <- "5"
  data[ data$Analysis == ID , "type"] <- "sample"
  
  for (ID in c(18364:18359)){
  ID2 <- ID-1
  data[ data$Analysis == ID , "Identifier1"] <- data[ data$Analysis == ID2 , "Identifier1"]
  data[ data$Analysis == ID , "mass"] <- data[ data$Analysis == ID2 , "mass"]
  data[ data$Analysis == ID , "carousel"] <- data[ data$Analysis == ID2 , "carousel"]
  data[ data$Analysis == ID , "type"] <- data[ data$Analysis == ID2 , "type"]
  }
  ID = 18359
  data[ data$Analysis == ID , "Identifier1"] <- "72h OD0.49"
  data[ data$Analysis == ID , "mass"] <- "17"
  data[ data$Analysis == ID , "carousel"] <- "5"
  data[ data$Analysis == ID , "type"] <- "sample"
  
  data$mass <- as.numeric(data$mass)
  
```

```{r define-stnds}
corr.std<-"act1"    #you should type the name of whichever standard you used for both your lin.std and drift.std here
C.acc <- -29.53
C.per.acc <- 71.09

dis.std<-"pugel"
C.acc.dis <- -12.6
C.per.dis <- 44.02

mon.std<-"EDTA2"
C.acc.mon <- -40.38
C.per.mon <- 41.09

# data frame overview
C.stds.table <- 
  tribble( # data frame defined by row, instead of by column
    ~std.name, ~C.acc, ~C.per.acc, # define the column names
    corr.std, C.acc, C.per.acc,   # add row one
    mon.std, C.acc.mon, C.per.mon, # add row two
    dis.std, C.acc.dis, C.per.dis  # add row three, etc.
  )

```

## Initial dataset check plots and culling
Plot yields of everything, and then just the standards, using INTERACTIVE plots. Use this to cull more samples if need be, by looking for  statistical outliers that coincide with yield problems.
Note that the linear model in the yield.stds figure here is based on ALL the standard data, so ALL stds that fall far off the line should be excluded from corrections

Check to see that samples all fall within linearity range, and for any other obvious outliers
to check:

* do linearity standards cover the area space of your samples and other standards?
* do all the standards show typical trends between area and mass? Do you see very general trends like that in your samples?
* are blanks actually blank? 
    * If not, but still small, may need to do a blank correction for carbon (is already a part of the nitrogen corrections) - if carbon is diluted, likely won't need to do a blank correction
    * If not, and the areas are big, check your raw data and original peaks to be sure there isn't something wrong with your input data files (e.g. offset lines, etc) or that the analyses were off during the run (e.g. offset accidently in carousel, so names go with incorrect sample data)

```{r cull.outliers.2, fig.width=10, fig.height=6}
# adjust next line only: change #'s in stds.to.cull to reflect the row #'s that need to be culled, add row#'s as needed and rerun after looking at the new plots
stds.all<-subset(data, type!="empty" & type!="sample")

stds.to.cull <- c(18306) #Analysis with yield problem and/or significant outlier in d13C or d18O, can rerun this line after 1st round to see outiers, than add them here   

stds.culled <- filter(stds.all, Analysis %in% stds.to.cull)
stds <- filter(stds.all, !Analysis %in% stds.to.cull) 

stds<- data.frame(stds, cbind(predict.lm(lm(stds$area44 ~ stds$mass), interval=c("confidence"), level=0.95)))

yield.C.all<-ggplot(data, aes(x=as.numeric(mass), y=area44, fill=type, label=Analysis)) +
  geom_point(size=3, shape=22) +
  theme_bw() +
  labs(title= "all C data")

yield.C.stds<-ggplot(subset(stds, type=="lin.std"), aes(x=as.numeric(mass), y=area44, label=Analysis)) +
  stat_smooth(method="lm") +   
  geom_point(data=stds, aes(fill=type, shape=type), size=2) +
  theme_bw() +
  scale_shape_manual(values=c(21,22,23,24,25,21,22,23)) +
  labs(title= "C standards - yield")

calc_std_means_d13C <- function(df) calc_means(df, "d13C")

d13C.stds <- 
  ggplot(stds, aes(label=Analysis)) +
  geom_hline(
    data = calc_std_means_d13C,
    mapping = aes(yintercept = yintercept, color = type, linetype = linetype)) +
  geom_point(shape=21, mapping = aes(x=area44, y=d13C, fill=type)) +
  scale_linetype_manual(values = c(1, 3, 2, 3, 2)) + 
  facet_grid(type ~ ., scales = "free") +
  theme_bw() +
  labs(title= "d13C standards - means and uncertainties")

ggplotly(yield.C.all) 
ggplotly(yield.C.stds)
ggplotly(d13C.stds)
```

```{r}
iso_plot_continuous_flow_data(dxf_files$`18336__blank_0.dxf`, c(44,45,46)) #what happend? left from last run?
```
```{r}
#iso_plot_continuous_flow_data(dxf_files$`16228__collagen_72.dxf`  , c(44,45,46)) #odd isotope value
```

## Carbon corrections
Apply basic offset correction to all of the data; does NOT include linearity or drift correction to data - culling any additional standards that should be culled (e.g. 2 sd outliers)
```{r offset.C, fig.width=10, fig.height=5}

offsetC<-subset(stds, Identifier1==corr.std)

(offsetC.mean<-mean(offsetC$d13C))
(offsetC.sd<-sd(offsetC$d13C))

offsetC$d13C.offset <- offsetC$d13C +  (C.acc - offsetC.mean)

(offsetcorrC.mean<-mean(offsetC$d13C.offset))
(offsetcorrC.sd<-sd(offsetC$d13C.offset))

d13C.offset<-ggplot(offsetC, aes(x=area44, y=d13C.offset, shape=type)) +
  geom_point(fill="orange", size=3) +
  geom_hline(yintercept=offsetcorrC.mean, colour="orange") +
  geom_hline(yintercept=offsetcorrC.mean + offsetcorrC.sd, colour="orange", linetype="dashed") +
  geom_hline(yintercept=offsetcorrC.mean - offsetcorrC.sd, colour="orange", linetype="dashed") +
  geom_hline(yintercept=offsetcorrC.mean + 2*offsetcorrC.sd, colour="orange", linetype=3) +
  geom_hline(yintercept=offsetcorrC.mean - 2*offsetcorrC.sd, colour="orange", linetype=3) +
  scale_shape_manual(values=c(21,22,23,24,25)) +
  annotate("text", y = offsetcorrC.mean + 0.01, x = min(offsetC$area44), 
    label = paste0("mean: ", sprintf("%.2f", offsetcorrC.mean), " \U00B1 ", sprintf("%.2f", offsetcorrC.sd, 2), " \U2030 (1 sd)"),
    size = 4, hjust=0, vjust=0, parse=FALSE, colour="black") +
  theme_bw() 

d13C.offset.mass<-ggplot (subset(stds, type=="lin.std"), aes(x=area44, y=mass)) +
  stat_smooth(method="lm") +
  geom_point(data=offsetC, aes(x=area44, y=mass), shape=21, fill="orange", size = 2) +
  theme_bw()

multiplot(d13C.offset, d13C.offset.mass, cols=2)
```

After finalizing the offset correction, apply it to the whole dataset, and check the monitoring standards
```{r apply.offset.corr.C, fig.width=10, fig.height=5}

#apply offset correction to whole dataset
data$d13C.offset <- data$d13C +  (C.acc - offsetC.mean)
stds$d13C.offset <- stds$d13C +  (C.acc - offsetC.mean)

#make monitoring standard dataset and dataset for additional standards used later for discrimination correction
offsetC.mon <- subset(stds, Identifier1==mon.std)
offsetC.dis <- subset(stds, Identifier1==dis.std)

#check monitoring standard response
(offsetC.mon.mean<-mean(offsetC.mon$d13C.offset))
(offsetC.mon.sd<-sd(offsetC.mon$d13C.offset))

C.stds.table$offsetC.mean <- c(offsetcorrC.mean, offsetC.mon.mean, mean(offsetC.dis$d13C.offset))
C.stds.table$offsetC.sd <- c(offsetcorrC.sd, offsetC.mon.sd, sd(offsetC.dis$d13C.offset))

C.mon.offset<-ggplot(offsetC.mon, aes(x=area44, y=d13C.offset)) +
  geom_point(shape=21, fill="orange") +
  geom_hline(yintercept=offsetC.mon.mean, colour="orange") +
  geom_hline(yintercept=offsetC.mon.mean + offsetC.mon.sd, colour="orange", linetype="dashed") +
  geom_hline(yintercept=offsetC.mon.mean - offsetC.mon.sd, colour="orange", linetype="dashed") +
  geom_hline(yintercept=offsetC.mon.mean + 2*offsetC.mon.sd, colour="orange", linetype=3) +
  geom_hline(yintercept=offsetC.mon.mean - 2*offsetC.mon.sd, colour="orange", linetype=3) +
  annotate("text", y = offsetC.mon.mean +0.01, x = min(offsetC.mon$area44), label = paste0("mean: ", sprintf("%.2f", offsetC.mon.mean), " \U00B1 ", sprintf("%.2f", offsetC.mon.sd, 2), " \U2030 (1 sd)"), size = 4, hjust=0, vjust=0, parse=FALSE) +
  theme_bw()

C.mon.offset.mass<-ggplot (subset(stds, type=="lin.std"), aes(x=mass, y=area44)) +
  stat_smooth(method="lm") +
  geom_point(data=offsetC.mon, aes(x=mass, y=area44), shape=21, fill="orange", size = 2) +
  theme_bw()

multiplot(C.mon.offset, C.mon.offset.mass, cols=2)
```

drift corrections, using raw values
```{r drift.C, fig.width=10, fig.height=5}
driftC<-subset(stds, type=="drift.std")

drift.slopeC<-(coef(lm(driftC$d13C ~ driftC$row))[[2]])
drift.interC<-(coef(lm(driftC$d13C ~ driftC$row))[[1]])

#drift check
driftC$d13C.drift<- driftC$d13C + (C.acc - (drift.slopeC * driftC$row + drift.interC))

(driftC.mean<-mean(driftC$d13C.drift))
(driftC.sd<-sd(driftC$d13C.drift))

C.drift<-ggplot(driftC, aes(x=row, y=d13C)) +
  geom_smooth(method=lm, colour="black") +
  annotate("text", x = min(driftC$row), y = max(driftC$d13C + 0.01), label = lm_eqn(driftC$row, driftC$d13C),  size = 4, hjust=0, vjust=0, parse=TRUE, colour="black") +
  geom_point(shape=21, fill="black", size=2) +
  geom_point(aes(x=row, y=d13C.drift), fill="red", shape=21, size=2) +
  geom_hline(aes(yintercept=C.acc), size=.5) +
  geom_hline(yintercept = driftC.mean + driftC.sd, colour="red", linetype="dashed") +
  geom_hline(yintercept = driftC.mean - driftC.sd, colour="red", linetype="dashed") +
  geom_hline(yintercept = driftC.mean + 2*driftC.sd, colour="red", linetype=3) +
  geom_hline(yintercept = driftC.mean - 2*driftC.sd, colour="red", linetype=3) +
  annotate("text", 
    y = driftC.mean +0.01, 
    x = min(driftC$inv.area44),
    label = paste0("mean: ", sprintf("%.2f", driftC.mean), " \U00B1 ", sprintf("%.2f", driftC.sd, 2), " \U2030 (1 sd)"),
    size = 4, hjust=0, vjust=0, parse=FALSE) +
    theme_bw()

C.drift.mass<-ggplot(subset(stds, type=="lin.std"), aes(x=area44, y=mass)) +
  stat_smooth(method="lm") +
  geom_point(data=driftC, aes(x=area44, y=mass), shape=21, fill="red", size=3) +
  theme_bw()

multiplot(C.drift, C.drift.mass, cols=2)
```

apply drift correction to all of the data, check with monitoring standards
```{r apply.drift.corr.C, fig.width=10, fig.height=5}
data$d13C.drift <- data$d13C +  (C.acc - (drift.slopeC * data$row + drift.interC))
stds$d13C.drift <- stds$d13C +  (C.acc - (drift.slopeC * stds$row + drift.interC))

driftC.mon<-subset(stds, Identifier1==mon.std)
driftC.dis <- subset(stds, Identifier1==dis.std)

(driftC.mon.mean<-mean(driftC.mon$d13C.drift))
(driftC.mon.sd<-sd(driftC.mon$d13C.drift))

C.stds.table$driftC.mean <- c(driftC.mean, driftC.mon.mean, mean(driftC.dis$d13C.drift))
C.stds.table$driftC.sd <- c(driftC.sd, driftC.mon.sd, sd(driftC.dis$d13C.drift))

C.mon.drift<-ggplot(driftC.mon, aes(x=area44, y=d13C.drift)) +
  geom_point(shape=21, fill="red") +
  geom_hline(yintercept = driftC.mon.mean, colour="red") +
  geom_hline(yintercept = driftC.mon.mean + driftC.mon.sd, colour="red", linetype="dashed") +
  geom_hline(yintercept = driftC.mon.mean - driftC.mon.sd, colour="red", linetype="dashed") +
  geom_hline(yintercept = driftC.mon.mean + 2*driftC.mon.sd, colour="red", linetype=3) +
  geom_hline(yintercept = driftC.mon.mean - 2*driftC.mon.sd, colour="red", linetype=3) +
  annotate("text",
    y = driftC.mon.mean +0.01, 
    x = min(driftC.mon$area44),
    label = paste0("mean: ", sprintf("%.2f", driftC.mon.mean), " \U00B1 ", sprintf("%.2f", driftC.mon.sd, 2), " \U2030 (1 sd)"),
    size = 4, hjust=0, vjust=0, parse=FALSE, colour="red")

C.mon.drift.mass<-ggplot (subset(stds, type=="lin.std"), aes(x=mass, y=area44)) +
  stat_smooth(method="lm") +
  geom_point(data=driftC.mon, aes(x=mass, y=area44), shape=21, fill="red", size = 2) +
  theme_bw()

multiplot(C.mon.drift, C.mon.drift.mass, cols=2)
```

Plot linearity
```{r linearityC, fig.width=10, fig.height=5}
linC<-subset(stds, type=="lin.std") 

lin.slopeC<-(coef(lm(linC$d13C ~ linC$inv.area44))[[2]])
lin.interC<-(coef(lm(linC$d13C ~ linC$inv.area44))[[1]])

#linearity check
linC$d13C.lin<-linC$d13C + (C.acc - (lin.slopeC * linC$inv.area44 + lin.interC))

(linC.mean<-mean(linC$d13C.lin))
(linC.sd<-sd(linC$d13C.lin))

C.lin.area44<-ggplot(linC, aes(x=area44, y=d13C)) +
  geom_point(shape=21, fill="blue") +
  geom_smooth() +
  theme_bw()

C.lincorr.invarea<-ggplot(linC, aes(x=inv.area44, y=d13C)) +
  geom_smooth(method=lm) +
  annotate("text", x = min(linC$inv.area44), y = max(linC$d13C + 0.01), label = lm_eqn(linC$inv.area44, linC$d13C),  size = 4, hjust=0, vjust=0, parse=TRUE) +
  geom_point(shape=21, fill="black", size=2) +
  geom_point(aes(x=inv.area44, y=d13C.lin), fill="red") +
  geom_hline(aes(yintercept=C.acc), size=.5) +
  geom_hline(yintercept = linC.mean + linC.sd, colour="red", linetype="dashed") +
  geom_hline(yintercept = linC.mean - linC.sd, colour="red", linetype="dashed") +
  geom_hline(yintercept = linC.mean + 2*linC.sd, colour="red", linetype=3) +
  geom_hline(yintercept = linC.mean - 2*linC.sd, colour="red", linetype=3) +
  annotate("text",
    y = linC.mean +0.01, 
    x = min(linC$inv.area44),
    label = paste0("mean: ", sprintf("%.2f", linC.mean), " \U00B1 ", sprintf("%.2f", linC.sd, 2), " \U2030 (1 sd)"),
    size = 4, hjust=0, vjust=0, parse=FALSE) +
    theme_bw()

C.lin.mass<-ggplot (subset(stds, type=="lin.std"), aes(x=mass, y=area44)) +
  stat_smooth(method="lm") +
  geom_point(data=linC, aes(x=mass, y=area44), shape=21, fill="blue", size = 2) +
  theme_bw()

multiplot(C.lin.area44, C.lincorr.invarea, C.lin.mass, cols=3)
```

apply linearity correction + offset correction to all of the data
```{r apply.lin.corr.C, fig.width=10, fig.height=5}

data$d13C.lin <- data$d13C +  (C.acc - (lin.slopeC * data$inv.area44 + lin.interC))
stds$d13C.lin <- stds$d13C +  (C.acc - (lin.slopeC * stds$inv.area44 + lin.interC))

linC.mon<-subset(stds, Identifier1==mon.std)
linC.dis<-subset(stds, Identifier1==dis.std)

(linC.mon.mean <- mean(linC.mon$d13C.lin))
(linC.mon.sd<-sd(linC.mon$d13C.lin))

C.stds.table$linC.mean <- c(linC.mean, linC.mon.mean, mean(linC.dis$d13C.lin))
C.stds.table$linC.sd <- c(linC.sd, linC.mon.sd, sd(linC.dis$d13C.lin))

C.mon.lin<-ggplot(linC.mon, aes(x=area44, y=d13C.lin)) +
  geom_point(shape=21, fill="blue") +
  geom_hline(yintercept = linC.mon.mean, colour="blue") +
  geom_hline(yintercept = linC.mon.mean + linC.mon.sd, colour="blue", linetype="dashed") +
  geom_hline(yintercept = linC.mon.mean - linC.mon.sd, colour="blue", linetype="dashed") +
  geom_hline(yintercept = linC.mon.mean + 2*linC.mon.sd, colour="blue", linetype=3) +
  geom_hline(yintercept = linC.mon.mean - 2*linC.mon.sd, colour="blue", linetype=3) +
  annotate("text",
    y = linC.mon.mean +0.01, 
    x = min(linC.mon$area44),
    label = paste0("mean: ", sprintf("%.2f", linC.mon.mean), " \U00B1 ", sprintf("%.2f", linC.mon.sd, 2), " \U2030 (1 sd)"),
    size = 4, hjust=0, vjust=0, parse=FALSE, colour="blue")

C.mon.lin.mass<-ggplot (subset(stds, type=="lin.std"), aes(x=mass, y=area44)) +
  stat_smooth(method="lm") +
  geom_point(data=linC.mon, aes(x=mass, y=area44), shape=21, fill="blue", size = 2) +
  theme_bw()

multiplot(C.mon.lin, C.mon.lin.mass, cols=2)
```

drift corrections on the linearity corrected values
```{r lindrift.C, fig.width=10, fig.height=5}

lindriftC <- merge(driftC, data[c("Analysis", "d13C.lin")], by.x="Analysis", by.y="Analysis", all.x=TRUE, all.y=FALSE, sort=FALSE)

lindrift.slopeC<-(coef(lm(lindriftC$d13C.lin ~ lindriftC$row))[[2]])
lindrift.interC<-(coef(lm(lindriftC$d13C.lin ~ lindriftC$row))[[1]])

#drift check
lindriftC$d13C.lindrift<- lindriftC$d13C.lin + (C.acc - (lindrift.slopeC * lindriftC$row + lindrift.interC))

(lindriftC.mean<-mean(lindriftC$d13C.drift))
(lindriftC.sd<-sd(lindriftC$d13C.drift))

C.lindrift<-ggplot(lindriftC, aes(x=row, y=d13C.lin)) +
  geom_smooth(method=lm, colour="black") +
  annotate("text", x = min(lindriftC$row), y = max(lindriftC$d13C.lin + 0.01), label = lm_eqn(lindriftC$row, lindriftC$d13C.lin),  size = 4, hjust=0, vjust=0, parse=TRUE, colour="black") +
  geom_point(shape=21, fill="black", size=2) +
  geom_point(aes(x=row, y=d13C.lindrift), fill="red", shape=21, size=2) +
  geom_hline(aes(yintercept=C.acc), size=.5) +
  geom_hline(yintercept = lindriftC.mean + lindriftC.sd, colour="red", linetype="dashed") +
  geom_hline(yintercept = lindriftC.mean - lindriftC.sd, colour="red", linetype="dashed") +
  geom_hline(yintercept = lindriftC.mean + 2*lindriftC.sd, colour="red", linetype=3) +
  geom_hline(yintercept = lindriftC.mean - 2*lindriftC.sd, colour="red", linetype=3) +
  annotate("text",
    y = lindriftC.mean +0.01, 
    x = min(lindriftC$inv.area44),
    label = paste0("mean: ", sprintf("%.2f", lindriftC.mean), " \U00B1 ", sprintf("%.2f", lindriftC.sd, 2), " \U2030 (1 sd)"),
    size = 4, hjust=0, vjust=0, parse=FALSE) +
    theme_bw()

C.lindrift.mass<-ggplot(subset(stds, type=="lin.std"), aes(x=area44, y=mass)) +
  stat_smooth(method="lm") +
  geom_point(data=lindriftC, aes(x=area44, y=mass), shape=21, fill="red", size=3) +
  theme_bw()

multiplot(C.lindrift, C.lindrift.mass, cols=2)
```

apply drift correction to all of the linearirty corrected data, check with monitoring standards
```{r apply.lindrift.corr.C, fig.width=10, fig.height=5}
data$d13C.lindrift <- data$d13C.lin +  (C.acc - (lindrift.slopeC * data$row + lindrift.interC))
stds$d13C.lindrift <- stds$d13C.lin +  (C.acc - (lindrift.slopeC * stds$row + lindrift.interC))

lindriftC.mon<-subset(stds, Identifier1==mon.std)
lindriftC.dis<-subset(stds, Identifier1==dis.std)

(lindriftC.mon.mean<-mean(lindriftC.mon$d13C.lindrift))
(lindriftC.mon.sd<-sd(lindriftC.mon$d13C.lindrift))

C.stds.table$lindriftC.mean <- c(lindriftC.mean, lindriftC.mon.mean, mean(lindriftC.dis$d13C.lindrift))
C.stds.table$lindriftC.sd <- c(lindriftC.sd, lindriftC.mon.sd, sd(lindriftC.dis$d13C.lindrift))

C.mon.drift<-ggplot(lindriftC.mon, aes(x=area44, y=d13C.lindrift)) +
  geom_point(shape=21, fill="red") +
  geom_hline(yintercept = lindriftC.mon.mean, colour="red") +
  geom_hline(yintercept = lindriftC.mon.mean + lindriftC.mon.sd, colour="red", linetype="dashed") +
  geom_hline(yintercept = lindriftC.mon.mean - lindriftC.mon.sd, colour="red", linetype="dashed") +
  geom_hline(yintercept = lindriftC.mon.mean + 2*lindriftC.mon.sd, colour="red", linetype=3) +
  geom_hline(yintercept = lindriftC.mon.mean - 2*lindriftC.mon.sd, colour="red", linetype=3) +
  annotate("text",
    y = lindriftC.mon.mean +0.01, 
    x = min(lindriftC.mon$area44),
    label = paste0("mean: ", sprintf("%.2f", lindriftC.mon.mean), " \U00B1 ", sprintf("%.2f", lindriftC.mon.sd, 2), " \U2030 (1 sd)"),
    size = 4, hjust=0, vjust=0, parse=FALSE, colour="red")

C.mon.drift.mass<-ggplot (subset(stds, type=="lin.std"), aes(x=mass, y=area44)) +
  stat_smooth(method="lm") +
  geom_point(data=lindriftC.mon, aes(x=mass, y=area44), shape=21, fill="red", size = 2) +
  theme_bw()

multiplot(C.mon.drift, C.mon.drift.mass, cols=2)
```


d13C discrimination correction - First, need to pick which correction scheme you prefer, then update this with the correct data columns
```{r apply.disc.corr.C}

# replace the character string here with the correction column you want to use
col_for_disc <- "lindrift"  # options to substitute in here are: "offset" or "drift" or  "lin" or  "lindrift"

# resulting formulas 
col_in_C.stds.table <- paste0(col_for_disc, "C.mean")
col_in_data <- paste0("d13C.", col_for_disc)
col_for_disc_mean_eq <- list(col_for_disc = lazyeval::interp(~var, var = as.name(col_in_C.stds.table)))
correction_eq <- list(d13C.disc = lazyeval::interp(~disc.slopeC * var + disc.interC, var = as.name(col_in_data)))

# safety checks
if (!col_in_C.stds.table %in% names(C.stds.table)) stop("this column does not exist in C.stds.table: ", col_in_C.stds.table, call. = FALSE)
if (!col_in_data %in% names(data)) stop("this column does not exist in data or stds table: ", col_in_data, call. = FALSE)

# regression
m <- lm(C.acc ~ col_for_disc, data = mutate_(C.stds.table, .dots = col_for_disc_mean_eq))
disc.slopeC<-(coef(m)[[2]])
disc.interC<-(coef(m)[[1]])
R2 <- summary(m)$r.squared

# apply correction
data <- mutate_(data, .dots = correction_eq)
stds <- mutate_(stds, .dots = correction_eq)

C.disc.all <- 
  ggplot(C.stds.table, aes_string(x=col_in_C.stds.table, y="C.acc")) +
  geom_smooth(method="lm", color = "blue") +
  geom_point(data=data, aes_string(x=col_in_data, y="d13C.disc"), shape=23, fill="red", size = 2) +
  geom_point(shape=21, fill="blue", size = 4) +
  geom_text(
           x = min(C.stds.table[[col_in_C.stds.table]]), 
           y = max(C.stds.table$C.acc), 
           label = str_interp("C.acc = $[.2f]{slope} ${var} + $[.2f]{intercept} (R2: $[.4f]{R2})", 
                              list(slope = disc.slopeC, intercept = disc.interC, var = col_for_disc, R2 = R2)),
           size = 4, hjust=0, vjust=0, colour="blue") +
  labs(x = col_for_disc)

C.disc.all
```
```{r add_disc_to_stds.mean.table}
discC.mon<-subset(stds, Identifier1==mon.std)
discC.dis<-subset(stds, Identifier1==dis.std)
discC.corr<-subset(stds, Identifier1==corr.std)

(discC.mon.mean<-mean(discC.mon$d13C.disc))
(discC.mon.sd<-sd(discC.mon$d13C.disc))

C.stds.table$discC.mean <- c(mean(discC.corr$d13C.disc), discC.mon.mean, mean(discC.dis$d13C.disc))
C.stds.table$discC.sd <- c(sd(discC.corr$d13C.disc), discC.mon.sd, sd(discC.dis$d13C.disc))
```

Calculate k-factor linearity to calculate weight percent C
```{r calc_percentC}
perC.stds.cull <- c(18333)

perC.stds <- filter(stds, !Analysis %in% perC.stds.cull)
perC.lin <-subset(perC.stds, type=="lin.std")

perC.lin$kfactor<-C.per.acc * perC.lin$mass / perC.lin$area44

avg.kfactor.C<-mean(perC.lin$kfactor)

#check percent C calculation by applying back to correcting standards and check mean and sd
perC.lin$percent.C<-avg.kfactor.C * perC.lin$area44 / perC.lin$mass

(percentC.avg<-mean(perC.lin$percent.C))
(percentC.sd<-sd(perC.lin$percent.C))

#plot k-factor versus 1/area, to see linearity, check for outliers
perCvinv.area<-ggplot(perC.lin, aes(x=inv.area44, y=kfactor)) +
  geom_smooth(method=lm) +
  geom_point(size=3) +
  annotate("text", 
           x = min(perC.lin$inv.area44), 
           y = max(perC.lin$kfactor + 0.01), 
           label = lm_eqn(perC.lin$inv.area44, perC.lin$kfactor),  
           size = 4, hjust=0, vjust=0, parse=TRUE)

perCvarea<-ggplot(perC.lin, aes(x=area44, y=kfactor, label=Analysis)) +
  geom_smooth() +
  geom_point(size=3)
ggplotly(perCvarea)
multiplot(perCvinv.area,perCvarea, cols=2)

```

```{r apply_percentC}
data$percent.C<-avg.kfactor.C * data$area44 / data$mass
perC.stds$percent.C<-avg.kfactor.C * perC.stds$area44 / perC.stds$mass

#check effect on monitoring standard
perC.monitor.std<-subset(data, Identifier1==mon.std)

perC.monitor.std.mean<-mean(perC.monitor.std$percent.C)
perC.monitor.std.sd<-sd(perC.monitor.std$percent.C)

perC.monitor.std.mean
perC.monitor.std.sd
```

```{r calc_percentC_drift}

perC.drift <- subset(perC.stds, type=="drift.std")

drift.slope.perC<-(coef(lm(perC.drift$percent.C ~ perC.drift$Analysis))[[2]])
drift.intercept.perC<-(coef(lm(perC.drift$percent.C ~ perC.drift$Analysis))[[1]])

#drift check
perC.drift$perC.drift<- perC.drift$percent.C + (C.per.acc - (drift.slope.perC * perC.drift$row + drift.intercept.perC))

(perC.drift.mean<-mean(perC.drift$perC.drift))
(perC.drift.sd<-sd(perC.drift$perC.drift))

drift.perC<-ggplot(perC.drift, aes(x=row, y=percent.C)) +
  geom_point(shape=21, fill="blue", size=3) +
  geom_smooth(method="lm", color="blue", se=FALSE) +
  geom_point(aes(x=row, y=perC.drift), shape=21, fill="red") +
  geom_smooth(aes(x=row, y=perC.drift), method="lm", color="red", se=FALSE) +
  annotate("text", x = 12, 
           y = C.per.acc, 
           label = lm_eqn(perC.drift$row, perC.drift$percent.C),  size = 4, hjust=0, vjust=0, parse=TRUE, colour="blue")

drift.perC
```

```{r apply_percentC_drift}
data$perC.drift<- data$percent.C + (C.per.acc - (drift.slope.perC * data$row + drift.intercept.perC))
perC.stds$perC.drift<- perC.stds$percent.C + (C.per.acc - (drift.slope.perC * perC.stds$row + drift.intercept.perC))

#check effect on monitoring standard
perC.monitor.std.drift<-subset(data, Identifier1==mon.std)

perC.monitor.std.mean.drift<-mean(perC.monitor.std.drift$perC.drift)
perC.monitor.std.sd.drift<-sd(perC.monitor.std.drift$perC.drift)

perC.monitor.std.mean.drift
perC.monitor.std.sd.drift
```
## Fractional Abundance conversion
```{r}
#devtools::install_github("sebkopf/isotopia")
library(isotopia)
data <- data %>% mutate(
    d13C.lindrift_Fractional_abundance=
        data$d13C.lindrift %>% 
        delta(notation = "permil",ref_ratio =  get_standard(name = "VPDB")) %>%
        to_ab()%>% 
        as.numeric()
    )
```

```{r save.data}
add_ws_with_data <- function(wb, sheet, data) {
  addWorksheet(wb, sheet)
  writeData(wb, sheet=sheet, data)
  return(wb)
}


monstd <- subset(data, type == "mon.std")
max(monstd$area44)
sampledata <- subset(data, type == "sample")
max(monstd$area44) < sampledata$area44
sampledatabig <- mutate(sampledata, runagain = max(monstd$area44) < area44)
sampledatabig <- mutate(sampledatabig, rerunmass= 30/(percent.C/100))

wb <- createWorkbook("data") 
wb <- add_ws_with_data(wb, "offsetC", offsetC)
wb <- add_ws_with_data(wb, "driftC", driftC)
wb <- add_ws_with_data(wb, "linC", linC)
wb <- add_ws_with_data(wb, "lindriftC", lindriftC)
wb <- add_ws_with_data(wb, "C Stds means", C.stds.table)
wb <- add_ws_with_data(wb, "all stds used", stds)
wb <- add_ws_with_data(wb, "culled stds", stds.culled)
wb <- add_ws_with_data(wb, "all data", data)
wb <- add_ws_with_data(wb, "reruns", subset(sampledatabig,runagain))

saveWorkbook(wb, paste0(session, "_corrected_data.xlsx"), overwrite = TRUE)
```

```{r}
samples<-filter(data, type=="sample")

d13Cvd18O<-ggplot(samples, aes(x=Start, y=d13C.disc, fill=Analysis)) +
  geom_point(size=3, shape=22) +
  theme_bw() +
  labs(title="Run 1 samples only")
  
ggplotly()
```